---
title: "NYT Geo Data"
output: html_notebook
theme: journal
---

<!-- Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.  -->

```{r}
getwd()
```

## Graphing the NYT over time 

I'm switching to R for this section of the project. I like Python's mapping capabilities just fine, but I prefer R's. 

First, import the data created earlier and some libraries  

```{r}
library(lubridate)
library(tidyverse)
library(zoo)
```


```{r}
full_archive <-  
    read.csv('complete_nyt_archive.csv')

full_archive.clean <- 
    full_archive %>% 
    mutate(pub_date = as.Date(pub_date), 
           year = year(pub_date), 
           month = month(pub_date), 
           yearmon = format(as.Date(pub_date), "%Y_%m")) %>% 
    select(c(yearmon, pub_date, web_url, lead_paragraph, news_desk, section_name, word_count)) %>% 
    filter(word_count != 0)
    
```



```{r}

df <- read.csv('data/mapped_df.csv')

data <- 
    df %>% 
    select(c(geo, lat, lng, news_desk, word_count, pub_date)) %>% 
    mutate(pub_date = as.Date(pub_date), 
           lat = as.numeric(lat), 
           lng = as.numeric(lng))

str(data)
```

Remember our original archive was some 575,000 rows; we have 87,112 with identifiable geographic data in their bylines. (Although agian, this is a bit aggressive, we're not including things like "ABOARD THE PAPAL AIRLINER" which... sure maybe has some geographic data in it.)

```{r}
nrow(data) / 339870

```
Looks like it's about 25%. How many of those come from each year? 
```{r}
summary
```

```{r}
summary %>%  ggplot() + geom_bar(aes(x = yearmon, y = count), stat = 'identity') + 
theme(panel.grid = element_blank(), 
          axis.text.x = element_text(angle = 75, vjust= 1, hjust = 1))
```


```{r}

summary <- 
    full_archive.clean %>% 
    group_by(yearmon) %>% 
    summarise(count = n()) 
print(summary)
summary %>% 
    ggplot(aes(x = yearmon, y = count, group = 1)) + 
    geom_bar(fill = 'dodgerblue2', alpha = 2/5, stat = 'identity') + 
    geom_abline(intercept = 6000, slope = -40, lty = 2, lwd = .9) + 
    geom_vline(xintercept = '2017_01', lty = 2, color = 'firebrick4', alpha = .975, lwd = .8) + 
    geom_segment( aes(x = '2018_01', y = 5750, 
                      xend = '2017_02', yend = 5750), 
                  arrow = arrow(length = (unit(0.5, 'cm')))) + 
    annotate('text', x = '2018_07', y = 5750, label = 'Trump elected') +
    theme_bw() + 
    scale_x_discrete('Date', breaks = c('2015', '2016')) + 
    ggtitle("Total Number of NYT Articles per Month, Jan, 2015 - July 2020") + 
    xlab('Date') +
    ylab("Number of Articles") + 
    theme(panel.grid = element_blank(), 
          axis.text.x = element_text(angle = 45, vjust= .5, hjust = 1))

```


```{r}
drop <- data[which(is.na(data$word_count)), ] 
```

```{r}
nrow(table(data$geo[1:10]))
```

Just in case we need it:

```{r}
yearly_words <- 
    data %>% 
    mutate(news_desk = as.factor(news_desk)) %>% 
    group_by(year(pub_date)) %>% 
    summarise(words = sum(word_count), .groups = 'keep') 
```

Now let's check out some summary statistics 

```{r}
summary <- 
    data %>% 
    group_by(year(pub_date)) %>% 
    summarise(
              words = sum(word_count), 
              avg_length = mean(word_count),
              article_count = n(), 
              place_diversity = nrow(table(geo))) 
summary
```
Note that we had to drop a few rows from 2018, and that will affect both the total word count and the `avg_length` of each article. 

Let's do some plotting. 

```{r}


```



---
<!-- Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*. -->

<!-- When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).  -->

<!-- The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed. -->

