---
title: "R Notebook"
output: html_notebook
---

```{r}
library(geojsonio)
library(broom)
#install.packages('rgeos')
library(rgeos)
```
## Intro and setup 
```{r}
# install.packages('rdgal') # working with geojson not shapefile
# library(rgdal)

hex_map <- geojson_read("data/us_states_hexgrid.geojson", what = "sp")

library(tidyverse)
library(sp)

hex_map@data <- #clean name
    hex_map@data %>% 
    mutate(google_name = gsub(" \\(United States\\)", "", google_name))
```

```{r}
plot(hex_map)
```

Ok, looks like we're good on step one. Now we try ggplot

```{r}
hex_map.fortified <- 
    tidy(hex_map, region = "google_name") 
centers <- cbind.data.frame(data.frame(gCentroid(hex_map, byid=TRUE), id=hex_map@data$iso3166_2))

```

```{r}
ggplot() +
  geom_polygon(data = hex_map.fortified, aes( x = long, y = lat, group = group), fill="skyblue", color="white") +
  geom_text(data=centers, aes(x=x, y=y, label=id)) +
  theme_void() +
  coord_map()
```

Now we'll add our data. We need to merge articles in with states, for which we need geocoding. 

```{r}
state_counts <- read.csv("data/state_counts.csv")

state_counts
```

Join on state to get our histogram -- let's use word count 

```{r}
# bin word count
hex_map.fortified <-
hex_map.fortified %>%
left_join(., state_counts, by = c('id' = 'state'))


hex_map.fortified$bin <-
    cut(hex_map.fortified$word_count,
        breaks = c(0, 50327, 247816, 874825, 564594, Inf),
        labels = c("0-25%", "0-25%", "25-50%", "50-75%", "75%+"))

hex_map.fortified


    
```

```{r}
yearly_counts <- read.csv("data/yearly_state_counts.csv")

table(yearly_counts$state) %>%
```


```{r}
library(viridis)
hex_map.fortified %>% 
    ggplot() +
    geom_polygon(aes(fill = bin, x = long, y = lat, group = group), color = 'gray', alpha = .74) + 
    theme_void() + 
    coord_map() +
    geom_text(data = centers, aes(x = x, y = y, label = id)) + 
    scale_fill_viridis(option = 'viridis', 
                       discrete = T) + 
    ggtitle("How many Words has the New York Times Filed from Each State?") + 
    labs(fill = 'Word Count Percentile') + 
    theme(plot.background = element_rect(fill = "#f5f5f2", color = NA), 
          panel.background = element_rect(fill = "#f5f5f2", color = NA), 
          legend.background = element_rect(fill = "#f5f5f2", color = NA),
          legend.position = 'bottom', 
          plot.title = element_text(family = '', 
                                    face = 'bold',
                                    hjust = .5))
    
    

```

Looks good. What can we see over time? We need to add a year to the hex_map df 

```{r}
yc_2015 <- yearly_counts %>% filter(year  == 2015)
```

```{r}
hex_map.fortified 
```
## Plotting changes over time 
```{r}
library(tidyverse)
library(zoo)
library(gtools)
yearly_counts_2015 <- 
  yearly_counts %>% rename(yearly_word_count  = word_count) %>% filter(year == 2015) 

yearly_counts_2015 %>% mutate(bin = cut(yearly_word_count, 
                                          breaks = c(quantile(yearly_counts_2015$yearly_word_count, probs = seq(0, 1, by = 0.25))),
                                          labels = c("0-25%", "25-50%", "50-75%", "75%+"))) %>% 
  right_join(hex_map.fortified %>% select(-c(word_count, bin)), by = c('state' = 'id')) %>% 
  na.fill0(0) %>% 
  ggplot() +
    geom_polygon(aes(fill = bin, x = long, y = lat, group = group), color = 'gray', alpha = .74) + 
    theme_void() + 
    coord_map() +
    geom_text(data = centers, aes(x = x, y = y, label = id)) + 
    scale_fill_viridis(option = 'viridis', 
                       discrete = T) + 
    ggtitle("How Much Space did the New York Times Give Each State in 2015?") + 
    labs(fill = 'Word Count Percentile') + 
    theme(plot.background = element_rect(fill = "#f5f5f2", color = NA), 
          panel.background = element_rect(fill = "#f5f5f2", color = NA), 
          legend.background = element_rect(fill = "#f5f5f2", color = NA),
          legend.position = "bottom", 
          plot.title = element_text(family = '', 
                                    face = 'bold',
                                    hjust = .5))
  
  
  
```

## making plots for each year 

```{r}
yearly_counts_2016 <- 
  yearly_counts %>% rename(yearly_word_count  = word_count) %>% filter(year == 2016) 

yearly_counts_2016 %>% mutate(bin = cut(yearly_word_count, 
                                          breaks = c(quantile(yearly_counts_2016$yearly_word_count, probs = seq(0, 1, by = 0.25))),
                                          labels = c("0-25%", "25-50%", "50-75%", "75%+"))) %>% 
  right_join(hex_map.fortified %>% select(-c(word_count, bin)), by = c('state' = 'id')) %>% 
  na.fill0(0) %>% 
  ggplot() +
    geom_polygon(aes(fill = bin, x = long, y = lat, group = group), color = 'gray', alpha = .74) + 
    theme_void() + 
    coord_map() +
    geom_text(data = centers, aes(x = x, y = y, label = id)) + 
    scale_fill_viridis(option = 'viridis', 
                       discrete = T) + 
    ggtitle("How Much Space did the New York Times Give Each State in 2016?") + 
    labs(fill = 'Word Count Percentile') + 
    theme(plot.background = element_rect(fill = "#f5f5f2", color = NA), 
          panel.background = element_rect(fill = "#f5f5f2", color = NA), 
          legend.background = element_rect(fill = "#f5f5f2", color = NA),
          legend.position = c(.525, .941), 
          legend.direction = 'horizontal',
          plot.title = element_text(family = '', 
                                    face = 'bold',
                                    hjust = .5))
    
```



```{r}
yearly_counts_2017 <- 
  yearly_counts %>% rename(yearly_word_count  = word_count) %>% filter(year == 2017) 

yearly_counts_2017 %>% mutate(bin = cut(yearly_word_count, 
                                          breaks = c(quantile(yearly_counts_2017$yearly_word_count, probs = seq(0, 1, by = 0.25))),
                                          labels = c("0-25%", "25-50%", "50-75%", "75%+"))) %>% 
  right_join(hex_map.fortified %>% select(-c(word_count, bin)), by = c('state' = 'id')) %>% 
  na.fill0(0) %>% 
  ggplot() +
    geom_polygon(aes(fill = bin, x = long, y = lat, group = group), color = 'gray', alpha = .74) + 
    theme_void() + 
    coord_map() +
    geom_text(data = centers, aes(x = x, y = y, label = id)) + 
    scale_fill_viridis(option = 'viridis', 
                       discrete = T) + 
    ggtitle("How Much Space did the New York Times Give Each State in 2017?") + 
    labs(fill = 'Word Count Percentile') + 
    theme(plot.background = element_rect(fill = "#f5f5f2", color = NA), 
          panel.background = element_rect(fill = "#f5f5f2", color = NA), 
          legend.background = element_rect(fill = "#f5f5f2", color = NA),
          legend.position = c(.525, .941), 
          legend.direction = 'horizontal',
          plot.title = element_text(family = '', 
                                    face = 'bold',
                                    hjust = .5))
```


```{r}
yearly_counts_2018 <- 
  yearly_counts %>% rename(yearly_word_count  = word_count) %>% filter(year == 2018) 

yearly_counts_2018 %>% mutate(bin = cut(yearly_word_count, 
                                          breaks = c(quantile(yearly_counts_2018$yearly_word_count, probs = seq(0, 1, by = 0.25))),
                                          labels = c("0-25%", "25-50%", "50-75%", "75%+"))) %>% 
  right_join(hex_map.fortified %>% select(-c(word_count, bin)), by = c('state' = 'id')) %>% 
  na.fill0(0) %>% 
  ggplot() +
    geom_polygon(aes(fill = bin, x = long, y = lat, group = group), color = 'gray', alpha = .74) + 
    theme_void() + 
    coord_map() +
    geom_text(data = centers, aes(x = x, y = y, label = id)) + 
    scale_fill_viridis(option = 'viridis', 
                       discrete = T) + 
    ggtitle("How Much Space did the New York Times Give Each State in 2018?") + 
    labs(fill = 'Word Count Percentile') + 
    theme(plot.background = element_rect(fill = "#f5f5f2", color = NA), 
          panel.background = element_rect(fill = "#f5f5f2", color = NA), 
          legend.background = element_rect(fill = "#f5f5f2", color = NA),
          legend.position = c(.525, .941), 
          legend.direction = 'horizontal',
          plot.title = element_text(family = '', 
                                    face = 'bold',
                                    hjust = .5))
```

```{r}
yearly_counts_2019 <- 
  yearly_counts %>% rename(yearly_word_count  = word_count) %>% filter(year == 2019) 

yearly_counts_2019 %>% mutate(bin = cut(yearly_word_count, 
                                          breaks = c(quantile(yearly_counts_2019$yearly_word_count, probs = seq(0, 1, by = 0.25))),
                                          labels = c("0-25%", "25-50%", "50-75%", "75%+"))) %>% 
  right_join(hex_map.fortified %>% select(-c(word_count, bin)), by = c('state' = 'id')) %>% 
  na.fill0(0) %>% 
  ggplot() +
    geom_polygon(aes(fill = bin, x = long, y = lat, group = group), color = 'gray', alpha = .74) + 
    theme_void() + 
    coord_map() +
    geom_text(data = centers, aes(x = x, y = y, label = id)) + 
    scale_fill_viridis(option = 'viridis', 
                       discrete = T) + 
    ggtitle("How Much Space did the New York Times Give Each State in 2019?") + 
    labs(fill = 'Word Count Percentile') + 
    theme(plot.background = element_rect(fill = "#f5f5f2", color = NA), 
          panel.background = element_rect(fill = "#f5f5f2", color = NA), 
          legend.background = element_rect(fill = "#f5f5f2", color = NA),
          legend.position = 'bottom', 
          plot.title = element_text(family = '', 
                                    face = 'bold',
                                    hjust = .5))
```

```{r}
yearly_counts_2020 <- 
  yearly_counts %>% rename(yearly_word_count  = word_count) %>% filter(year == 2020) 

yearly_counts_2020 %>% mutate(bin = cut(yearly_word_count, 
                                          breaks = c(quantile(yearly_counts_2020$yearly_word_count, probs = seq(0, 1, by = 0.25))),
                                          labels = c("0-25%", "25-50%", "50-75%", "75%+"))) %>% 
  right_join(hex_map.fortified %>% select(-c(word_count, bin)), by = c('state' = 'id')) %>% 
  na.fill0(0) %>% 
  ggplot() +
    geom_polygon(aes(fill = bin, x = long, y = lat, group = group), color = 'gray', alpha = .74) + 
    theme_void() + 
    coord_map() +
    geom_text(data = centers, aes(x = x, y = y, label = id)) + 
    scale_fill_viridis(option = 'viridis', 
                       discrete = T) + 
    ggtitle("How Much Space did the New York Times Give Each State in 2020?") + 
    labs(fill = 'Word Count Percentile') + 
    theme(plot.background = element_rect(fill = "#f5f5f2", color = NA), 
          panel.background = element_rect(fill = "#f5f5f2", color = NA), 
          legend.background = element_rect(fill = "#f5f5f2", color = NA),
          legend.position = 'bottom', 
          plot.title = element_text(family = '', 
                                    face = 'bold',
                                    hjust = .5))
```

Ok so let's where the biggest shifts occurred 





## Two more state-by-state-maps: 
- Words per capita 
- Subtract election coverage and see what we have 

### get data 
```{r}
state_pop <- 
  read.csv('data/state_pop.csv') %>% 
  rename(pop_2010 = X2010) %>% drop_na() %>% mutate(estimate_2019 = as.numeric(estimate_2019))

state_pop 

```

```{r}
hex_map.fortified %>% left_join(state_pop, by = c('id' = 'State')) %>% 
  mutate(words_per_million_pop = word_count/((estimate_2019) / 1e6))
```

### plot 
```{r}
hex_map.fortified %>% left_join(state_pop, by = c('id' = 'State')) %>% 
  mutate(words_per_million_pop = (word_count/ 1e6) / ((estimate_2019) / 1e6)) %>%
  filter(id != "District of Columbia") %>% 
    ggplot() +
    geom_polygon(aes(fill = words_per_million_pop, x = long, y = lat, group = group), color = 'gray', alpha = .74) + 
  geom_polygon(data = hex_map.fortified %>% filter(id == 'District of Columbia'), aes(x = long, y = lat), fill = 'gray', alpha = .75) + 
    theme_void() + 
    coord_map() +
    geom_text(data = centers, aes(x = x, y = y, label = id)) + 
    scale_fill_viridis(option = 'viridis', 
                       discrete = F) + 
    ggtitle("How Many Words per Person has the New York Times Filed since 2015?") + 
    labs(fill = 'Words per Person') + 
    theme(plot.background = element_rect(fill = "#f5f5f2", color = NA), 
          panel.background = element_rect(fill = "#f5f5f2", color = NA), 
          legend.background = element_rect(fill = "#f5f5f2", color = NA),
          legend.text = element_text(face = 'bold'),
          legend.title = element_text(face = 'bold', 
                                      vjust = .8),
          legend.position = c(.5, .91), 
          legend.direction = 'horizontal',
          plot.title = element_text(family = '', 
                                    face = 'bold',
                                    hjust = .5))
```

### without politics coverage 

```{r}

state_counts_no_politics <-  
  read.csv('data/state_counts_no_politics.csv')
state_counts_no_politics


```


```{r}
hex_map.fortified %>% 
  select(-c(word_count, bin)) %>% 
  left_join(state_counts_no_politics, by = c("id" = "state")) %>% 
  left_join(state_pop, by = c('id' = 'State')) %>% 
  mutate(words_per_million_pop = (word_count/ 1e6) / ((estimate_2019) / 1e6)) %>%
  filter(id != "District of Columbia") %>%
    ggplot() +
    geom_polygon(aes(fill = words_per_million_pop, x = long, y = lat, group = group), color = 'gray', alpha = .74) +
  geom_polygon(data = hex_map.fortified %>% filter(id == 'District of Columbia'), aes(x = long, y = lat), fill = 'gray', alpha = .75) +
    theme_void() +
    coord_map() +
    geom_text(data = centers, aes(x = x, y = y, label = id)) +
    scale_fill_viridis(option = 'viridis',
                       discrete = F) +
    ggtitle("How Many Words per Person has the New York Times Filed since 2015?") +
    labs(fill = 'Words per Person') +
    theme(plot.background = element_rect(fill = "#f5f5f2", color = NA),
          panel.background = element_rect(fill = "#f5f5f2", color = NA),
          legend.background = element_rect(fill = "#f5f5f2", color = NA),
          legend.text = element_text(face = 'bold'),
          legend.title = element_text(face = 'bold',
                                      vjust = .8),
          legend.position = c(.5, .91),
          legend.direction = 'horizontal',
          plot.title = element_text(family = '',
                                    face = 'bold',
                                    hjust = .5))
```





